---
title: "UMD Fees"
---

Follow along with the demo to learn how to calculate percentage change using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.


### Task 1: Load the tidyverse


```{r}
library(tidyverse)
```




### Task 2: Load the UMD Fees dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv

```{r}
umd_fees <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv")
```

### Task 3: Calculate the percentage change in each category between Fall 2026 and Fall 2021 using `mutate()`

```{r}
umd_fees_pct <- umd_fees |> mutate(pct_change = (`Fall 2026`- `Fall 2021`)/`Fall 2021` * 100)
```



### Task 4: Describe the results

Write a few sentences about what you found in those percentages. What's the story here? What questions do you have?

Student sustainability had the largest change, increasing by 150% from Fall 2021 to Fall 2026. Looking at SustainableUMD, it says "Addressing sustainability is essential because healthy water, air, and ecosystems are foundational to a thriving society." Based on the graphic, it appears that the university has stepped-up its efforts in having a healthy environment. 

Another stat that stood out was the 58% increase in shuttle bus fees. Walking around campus, I have seen more electric buses so that may have something to do with the increase. There is also a notable increase from Fall 2023 to Fall 2024 (16%); that is interesting. 

One other note is the fact that athletics hasn't changed once. Who knows why it's 399 and not 400 exactly. 



When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
